Influence of Connectivity on Dung Beetle Communities

### 2025-06-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
clean up doc

# date: "\\vspace{-1em}"
header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
 - \usepackage{lineno}
# - \linenumbers
 - \usepackage[default]{sourcesanspro}
 - \usepackage{fancyhdr} # modify header into left, center, right
 - \pagestyle{fancy} # cues the use of fancy headers
 - \fancyfoot{} # cues the use of fancy footer
 - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
 - \fancyhead[R]{p. \thepage} # name and page on header right
# # - \fancyfoot[R]{ }
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output:
  pdf_document:
    includes:
      # in_header: style/preamble.tex
      template: null
      keep_tex: no
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

@@

1) Are any of the Dung Beetles we collected considered "open habitat" specialists? (might want to look at the work on pastures in the southeast)

cvig - open sand chaparral, open grassland preference nealis 1977
mbis - open sand chapparal preference nealis 1977
open - slight preference towards open chap and shaded chap, little open grassland nealis 1977

pign - sharing forest habitat but need more info conover 2019
pvin - open habitat but share forest with pign conover 2019
alec - forest preference conover 2019

2) Is there anything about these species that might be interesting in the context of corridors?
  - are any species known to be good fliers/movers?
    - Ospina 2018 -> wing shape py habitat preference and differences in wing shape determined by species groupings
    ^ beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies)
    ^ large body beetles tending towards lower energy flight strategies. maybe justification for corridor preference?

    = stanbrook and king 2022 -> tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.

  - specialists on deer dung?
    - Conover 2019 -> dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
    - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used. However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.

  - navigate by starlight?
  - feeding preferences?

@@

4) What are hill numbers?

hill numbers are a set of biodiversity metrics designed to standardize metrics and quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)

5) What is the advantage of useing hill numbers to estimate diversity / richness?

- Chao et al. 2014 -> incorporating abundance into calculating diversity metrics is important for studying functionality in particular since some species that may not be as abundant as others may not contribute to an ecosystem function as much as others.

- ^ At the same time rare species may be filling a specific niche or we might be concerned about exotics or at risk species.

@@

### Methods: 
1. ~~add a few sentences on drying, keys used to ID, and depositing vouchers~~

### Analyses
1. ~~Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm). ~~
1. Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)

### Results
1. answer the overview questions, them stich them together into a paragraph
 - ~~any species restricted to 1-2 habitat types~~
 - ~~Were all species found in all blocks? Were any restricted to only 1-2 blocks?~~
 - Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
 - Number of species in each functional group
 - ~~most common 3-4 species~~
 - any rare species? 
 - any invasive / exotic species?

@@

**_Note from EB:_** _I don't want to see the words 'dung beetles' until paragraph 5_.

Paragraph 1: Overall Topic & Why it is interesting and Important. _Fragmentation is an extensive problem across the globe and it is vital to understand shifts in biodiversity and community structures as a result of trophic cascades._

>As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented. Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018]. However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012]. This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].

Paragraph 2: What we know. _Habitat connectivity is generally viewed as beneficial for ecosystems as a whole. We know that patch shape and size can be an influental factor for how populations are configured. Wildlife corridors_

>Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995). Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity.

Paragraph 3: What we don't know. _Changes in ecosystem fuction as a result of shifts in community structure (EB but we may not want to emphasize this given we didn't measure directly_

>By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes. Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected. Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services.

@@

Paragraph 4: Study System and why it is so excellent for this issue. _Here we talk about dung beetles how they fit into the broader scope of our question and directly state what questions we are asking._

>Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked are corridors benefiting any one functional trait over another?

@@

Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals (Tewksbury et al. 2002). Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna. In each replicant the central patch (100 x 100 m) is always connected to one peripheral patch with identical dimensions by a 150 x 25 m corridor, this will hereafter be referred  to as the connected patch. The remaining patches are either "winged" or "rectangular".  The winged patch is also 100 x 100 m, however they exhibit their characteristic wings in the form of two 75 x 25 m offshoots meant to account for the extra area and edge space the corridor provides. The rectangular patch is 100 x 137.5 m also the same area as the space of the connected patch plus the corridor. Each block has a duplicate of either the winged or rectangle patch, all peripheral patches being 150 m from the center patch. For this study sampling was done in one of each patch type and in one matrix plot per block, all matrix blocks were set up 150 m away from the center as well.

@@

## Analysis

Biodiversity between patch types was compared using Hill numbers (Jost 2006). Hill numbers are metrics developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity. [delete: In our work, we use Hill numbers to explore the impacts of abundance in our diversity measurements]. They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation. [maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity). [can you add a sentence about how these are calculated? by whihc I mean - are the calculations different than for richness, shannon, etc.?] Diversity numbers and species richness were calculated in R studio using package `hill`.

To test for the effects of connectivity on species abundance, richness, and diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix]. For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution. We looked at [compared...looked at sounds squishy. try to use more active verbs like compared, quantified, assessed, etc.] (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type. We included the identity of the sampling block as a random effects. To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution. All models were fit using `lme4` package.  Prior to conducting our modeling we used _qqplots_ (`DARMa` package) to evaluate the the suitability of our data.

### Overview

<!-- _Don't have to describe these, just do it and report in results. this is just to remind us of what to do_ -->

<!-- 1. how many beetles total   -->
<!-- 1. most common species overall.   -->
<!-- 1. was. the same species the most common one in all habitats?   -->
<!-- 1. were there any that were caught in all habitats? Any restricted to only one?   -->
<!-- 1. any invasive / exotic species?   -->
<!-- 1. Table: count by species by habitat.  -->

### Species richness

To Test for differences in species richness, we....

1. glmm with poisson dist reccomended by julian
1. table: results of analyses. 
1. visualization: hill's plots? rank-abundance plots? rarefaction curve?

### Species diversity

We used a similar approach to compare species diversity in winged, isolated, and connected patches....

1. glmm with poisson dist reccomended by julian
1. table: results of analyses
1. visualization: hill's plots? rarefaction curve?

### Functional diversity

1. 1nd pass: roller, tunneler, dweller
1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025)
1. table: results of analyses 
1. visualization: biomass of different (R/T/D) functional groups?

_Reason for this breakdown: _ - richness and diversity are "community level" response. abundance is a species-level response. changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to functional group.

---

> Biodiversity between patch types was compared using Hill numbers (Jost 2006). We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity). Diversity numbers were calculated in R studio using package iNEXT (Chao et al. 2016). Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio. Dung beetles were assigned traits by waste removal guild and habitat preference.

> Functional Diversity: Need to assign each species to a functional group: roller, tunneler. dweller, others?  - habitat preference (forest, pasture, generalist)
 - what stats / computer packages did you use to analyze and visualize the data?

>_Look through dung beetle pubs and see how/what people compare_

<!--   - lets hammer this out  -->
<!--   - modeling?  -->
<!--     - glmm with poisson dist reccomended by julian -->
<!--   - beta, abundance, biomass? per site -->
<!--   - ~species list by sampling blocks (anything with this?)~ -->
<!--   - habitat preference  -->
<!--   - rarefaction -->

@@

## Community Composition and Abundance

Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species. The N = `r how_many_common` species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: _Canthon vigilans_ (N = `r most_common_spp[1,1]`), _Ateuchus lecontei_ (N = `r most_common_spp[2,1]`), _Phanaeus igneus_ (N = `r most_common_spp[3,1]`), _Dichotomius carolinus_ (N = `r most_common_spp[4,1]`), _Aphodeus alloblackburneus_ (N = `r most_common_spp[5,1]`), and _Onthophagus pennsylvanicus_ (N = `r most_common_spp[6,1]`; Table 1). All but four species were captured in every patch type. _Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches. _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).

Wegen comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangulat patches. Statistical analysis of abundance relating to abundance focused on the six most abundant species.

@@

## Species Richness

@@

## Biodiversity

### Overview

1. total number of beetles from total number of species: `r total_N` beetles from `r total_spp` species.

1. All but four species species were found in every patch type. _Onthophagus concinnus_ was found in the matrix and winged patches, while _Onthophagus striatulus_ was found in matrix habitat and rectanular patches. _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).

1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?

1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)

1. Number of species in each functional group

1. most common 3-4 species: The `r how_many_common` most common species were \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), and \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`).
1. any rare species? 
1. any invasive / exotic species?

### Species richness

### Species diversity

### Functional diversity

@@

Recap main goals into findings
 abundances in matrix vs connected patch and why this could be happening
 - source pop to habiitat edge

 species richness again supporting that matrix is more of an ideal habitat for dung beetle community

 diversity indices community structures weren't highly different between patch types

 bray curtis hinting at similar land uses between corridor and winged patch
 - like julians paper corridors benefit certain populations and more fit populations are able to make better use

Interpretation of results
 Abundance 
 - reasin for highest abundances in matrix and connected patch
 - lower abundance in rectangle hinting at fragmentation effects
 Species Richness & Diversity
 - the role of hbitat connectivity in shaping community structure or lack thereof
 - why do we think connnected had lowest species richness??
 Functional traits and ecological impacts
 - did corridoors favor a functional trait
 - why might there be a trait response 
 - implications for ecosystem processes
 - like duung removal papers seed dispersal and yep

Comparing to previous studies
- how are things aligning
- think about the biology

Limitations and future work
- potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
- other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)

Takeaways for conservation and management
- dung beetles are robust
- what do think about corridor design and considerations for fragmented lanndscapes
- practical applications think about the beetles

>>>>

# date: "\\vspace{-1em}"
header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
 - \usepackage{lineno}
 - \linenumbers
 - \usepackage[default]{sourcesanspro}
 - \usepackage{fancyhdr} # modify header into left, center, right
 - \pagestyle{fancy} # cues the use of fancy headers
 - \fancyfoot{} # cues the use of fancy footer
 - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
 - \fancyhead[R]{p. \thepage} # name and page on header right
# # - \fancyfoot[R]{ }
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output:
  pdf_document:
    includes:
      # in_header: style/preamble.tex
      template: null
      keep_tex: no
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

@@

~~1) Are any of the Dung Beetles we collected considered "open habitat" specialists? (might want to look at the work on pastures in the southeast)~~

- cvig - open sand chaparral, open grassland preference nealis 1977
- mbis - open sand chapparal preference nealis 1977
- open - slight preference towards open chap and shaded chap, little open grassland nealis 1977

- pign - sharing forest habitat but need more info conover 2019
- pvin - open habitat but share forest with pign conover 2019
- alec - forest preference conover 2019

~~2) Is there anything about these species that might be interesting in the context of corridors? Are any species known to be good fliers/movers? Specialists on deer dung? navigate by starlight?~~

  - Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings. Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies). Large body beetles tending towards lower energy flight strategies. maybe justification for corridor preference?
    - stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal. 
    - Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
    - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used. However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.

@@

~~4) What are hill numbers?~~

- hill numbers are a set of biodiversity metrics designed to standardize metrics and quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)

~~5) What is the advantage of useing hill numbers to estimate diversity / richness?~~

- Chao et al. 2014: incorporating abundance into calculating diversity metrics is important for studying functionality in particular since some species that may not be as abundant as others may not contribute to an ecosystem function as much as others. At the same time rare species may be filling a specific niche or we might be concerned about exotics or at risk species.

@@

### Methods:

1. ~~add a few sentences on drying, keys used to ID, and depositing vouchers~~

### Analyses

1. ~~Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm).~~

1. Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)

### Results

1. answer the overview questions, them stich them together into a paragraph

 - ~~any species restricted to 1-2 habitat types~~
 - ~~Were all species found in all blocks? Were any restricted to only 1-2 blocks?~~
 - ~~Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)~~
 - Number of species in each functional group
 - ~~most common 3-4 species~~
 - any rare species? 
 - any invasive / exotic species?

@@

As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented. Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018]. However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012]. This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995). Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity.

By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes. Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected. Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services.

@@

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked are corridors benefiting any one functional trait over another?

@@

Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals (Tewksbury et al. 2002). Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna. In each block the central patch (100 x 100 m) is always connected to one peripheral patch with identical dimensions by a 150 x 25 m corridor, this will hereafter be referred  to as the connected patch. The remaining patches are either "winged" or "rectangular".  The winged patch is also 100 x 100 m, however they exhibit their characteristic wings in the form of two 75 x 25 m offshoots meant to account for the extra area and edge space the corridor provides. The rectangular patch is 100 x 137.5 m also the same area as the space of the connected patch plus the corridor. Each block has a duplicate of either the winged or rectangle patch, all peripheral patches being 150 m from the center patch. For this study sampling was done in one of each patch type and in one matrix plot per block, all matrix blocks were set up 150 m away from the center as well.

@@

## Analyses

Biodiversity between patch types was compared using Hill numbers (Jost 2006). Hill numbers are metrics developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity. [delete: In our work, we use Hill numbers to explore the impacts of abundance in our diversity measurements]. They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation. [maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity). _[can you add a sentence about how these are calculated? by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT (Chao et al. 2016). Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio. Dung beetles were assigned traits by waste removal guild and habitat preference.

To test for the effects of connectivity on  abundance, speciesrichness, and species diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix]. For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution. We looked at [compared...looked at sounds squishy. try to use more active verbs like compared, quantified, assessed, etc.] (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type. We included the identity of the sampling block as a random effects. To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution. All models were fit using `lme4` package.  Prior to conducting our modeling we used _qqplots_ (`DARMa` package) to evaluate the the suitability of our data.

<!-- ### Functional diversity -->

<!-- 1. 1nd pass: roller, tunneler, dweller -->
<!-- 1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025) -->
<!-- 1. table: results of analyses  -->
<!-- 1. visualization: biomass of different (R/T/D) functional groups?    -->

<!-- _Reason for this breakdown: _ - richness and diversity are "community level" response. abundance is a species-level response. changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to functional group.  -->

<!-- Functional Diversity: Need to assign each species to a functional group: roller, tunneler. dweller, others?  - habitat preference (forest, pasture, generalist). Look through dung beetle pubs and see how/what people compare_ what stats / computer packages did you use to analyze and visualize the data? -->

@@

## Overview: Community Composition and Abundance

Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species. The N = `r how_many_common` species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1). All but four species were captured in every patch type. _Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches. _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).

- any rare species? 
- any invasive / exotic species?

When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangulat patches. Statistical analysis of abundance relating to abundance focused on the six most abundant species.

@@

### Species Richness

@@

### Species Diversity

### Functional Diversity

@@

## Recap main goals into findings

 1. Abundances in matrix vs connected patch and why this could be happening
 - source pop to habiitat edge

2. species richness again supporting that matrix is more of an ideal habitat for dung beetle community

3. diversity indices community structures weren't highly different between patch types

4. bray curtis hinting at similar land uses between corridor and winged patch
 - like julians paper corridors benefit certain populations and more fit populations are able to make better use

## Interpretation of results

1. Abundance 
 - reason for highest abundances in matrix and connected patch
 - lower abundance in rectangle hinting at fragmentation effects
2. Species Richness & Diversity
 - the role of hbitat connectivity in shaping community structure or lack thereof
 - why do we think connnected had lowest species richness??
3. Functional traits and ecological impacts
 - did corridoors favor a functional trait
 - why might there be a trait response 
 - implications for ecosystem processes
 - like duung removal papers seed dispersal and yep
4. Comparing to previous studies
- how are things aligning
- think about the biology 
5. Limitations and future work
- potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
- other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)
6. Takeaways for conservation and management
- dung beetles are robust
- what do think about corridor design and considerations for fragmented lanndscapes
- practical applications think about the beetles
